Effects of the body force on the pedestrian and the evacuation dynamics

### 2020-03-04 paper.tex
hice word wrap 80 lineas en los archivos de texto

\begin{table}
\begin{tabular}{c@{\hspace{6mm}}c@{\hspace{6mm}}c@{\hspace{6mm}}c@{\hspace{6mm}}
c@{\hspace{14mm}}l}
 \hline
 $\tau$[s]   & $m$[kg]     & $v_d$[m/s]  &  $B$[m]  & $k_n$[kg/s$^2$] &  Refs. \\
 \hline
0.61         & ---         & 1.24 & $0.36+1.06\,v$ &  ---                 &  
 \cite{seyfried_2007} \\
0.50$^*$     & 80$^*$      & 1.34 & 0.50           &  ---                 &  
\cite{weidmann_1992,lakoba_2005}\\
---          & $67.5$      & 1.39 &  ---           &  $96.1 + 12694.1\,x$ & 
\cite{song_2019}\\
---          & $67.0$      & 1.39 &  ---           &  $97.0 + 29378.9\,x$ & 
\cite{song_2019}\\

@@

\subsection{The set of parameters}

@@

\subfloat[\label{venn_2parameters}]
{\includegraphics[width=0.49\columnwidth]{./venn_2parameters.png}}
\hfill    
\subfloat[\label{venn_3parameters}]
{\includegraphics[width=0.49\columnwidth]{./venn_3parameters.png}
}\\
\caption[width=0.47\columnwidth]{Venn diagrams for the dimensionless parameters
appearing in the equation of motion (\ref{eqn_motion}) (see 
Appendix~\ref{appendix1} for details). The sets correspond to 
$\mathcal{A}=\{\tau/m,1/v_d,A\}$, $\mathcal{K}=\{\tau/m,B,k_t\}$ and 
$\mathcal{K}_c=\{\tau/m,1/v_d,B,k_n\}$. (a) The Venn diagram representation if no 
body force is introduced in the SFM (only sets $\mathcal{A}$ and $\mathcal{K}$, 
as in Ref.~\cite{dorso_2019}). (b) The Venn diagram representation for the sets 
$\mathcal{A}$, $\mathcal{K}$ and $\mathcal{K}_c$. }
\label{venn_diagram}
\end{figure}

@@

\begin{itemize}
 \item[(a)] The ``faster-is-slower'' (or the ``faster-is-faster'') phenomenon 
occurs when varying the values of $v_d$ (see Section~\ref{bottleneck}). This is 
equivalent to vary simultaneously $\mathcal{A}$ and $\mathcal{K}_c$ by the same 
amount in the (dimensionless) parameter space (see Section~\ref{parameters}). We 
may visualize this sampling procedure as moving along a straight line in the 
space $(\mathcal{A},\mathcal{K},\mathcal{K}_c)$. Further variations of 
$k_n$ turns the sample points out of this line, but on the plane of constant
$\mathcal{K}_c$. We will step up $k_n$ by several orders of magnitude in order 
to get the big picture of this surface.

@@

\item[(c)] We will explore crowd densities allowing body compressions up to 
$0.1\,$m. This corresponds to remarkably high densities, since our own 
(real-life) estimates do not surpass $0.065\,$m (see 
Section~\ref{experimental}). Therefore, the extremely high density 
scenarios (say, above $0.065\,$m) will be considered for the purpose of a 
tendency, but not as a common real-life situation. No casualties due to high 
pressures will be further considered.

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./vd_vs_te_N225.eps}
\caption{\label{vd_vs_te}Mean evacuation time (s) vs. the pedestrian’s desired  
velocity (m/s) for a bottleneck. The room was 20 m x 20 m size. The door was 
0.92 m width (two pedestrians' width). Mean values were computed from 10 evacuation processes. 225 
pedestrians were initially placed in a square lattice with a random initial 
velocity. Each process was finished when 158 pedestrians left the room. The 
different symbols indicate the $k_n$ value corresponding to the body force (see 
the label). The crosses correspond to the Helbing's original SFM parameter, the 
up-triangles correspond to the value measured in 
Ref.~\cite{melvin1988aatd}, squares correspond 
to zero body force and circles correspond to an extreme value of stiffness (one 
order of magnitude higher than the original SFM). The 
 down triangles correspond to an intermediate value between the empirical value 
presented in Ref.~\cite{melvin1988aatd} and the one provided by Helbing in 
Ref.~\cite{helbing_2000} }
\end{figure}

Three behavioral patterns can be distinguished in Fig.~\ref{vd_vs_te}. 
Each pattern can display a positive slope, a negative slope
or both. The interval in which the slope is positive means that the 
harder the pedestrians try to get out (higher $v_d$), the 
longer it takes them to evacuate. This is the Faster-is-slower 
(FIS) regime. Conversely, the interval in which the slope is 
negative  corresponds to a Faster-is-Faster (FIF) regime (the 
harder they try, the quicker they leave).\\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the pedestrian’s desired velocity (m/s). (b) Mean overlap as a function of the pedestrians desired velocity. Each symbol indicates the $k_n$ value corresponding to the body force (see the labels). The data corresponds to a bottleneck with periodic boundary conditions (re-injecting pedestrians). The average was taken over time and the pedestrians in the bottleneck. The sampling was done every five seconds once the crowd reached the stationary state (say, $t=20\,$s) until the end of the simulation ($t=1000\,$s). Color online only.}
\label{degree_overlap_vd}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
        \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\
\caption[width=0.47\columnwidth]{Snapshots of the contact networks of a 225 pedestrian evacuation through a bottleneck. The door is placed at $(x,y)=(20,10)\,$m, the width of the door is 0.92$\,$m (equivalent to 2 pedestrian\textsc{\char13}s diameter). The lines that connect the nodes (pedestrians) represent the contact between them. The color represents the degree (the number of pedestrians with which it is connected). (a) and (b) correspond to a simulation without body force with $v_d=$2 and $v_d=$10 respectively. (c) and (d) correspond to simulations with $k_n=1.2\,$E6 with $v_d=$2 and $v_d=$10 respectively. Color online only.}
\label{network_bottleneck}
\end{figure}

@@

It is worth noting from Fig.~\ref{degree_dens} that ``stiff''
pedestrians exhibit a sharper transition to the maximum packing density than 
the ``soft'' pedestrians. We further checked this phenomenon by counting the 
number of loops of three nodes (say, the triangles) present in the network. 
This magnitude was reported to be a sensitive magnitude for characterizing the 
jamming transition in a compressed granular packing (see 
Ref.~\cite{pugnaloni_2013} for details).  As can be seen in 
Fig.~\ref{triangles}, the ``stiff'' pedestrians share more triangles 
than the ``soft'' ones.   \\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the corridor at time $t=50\,$s. The global density was $\rho=6$.  The lines that connect the nodes (pedestrians) represent the contact between them. The colors stand for the degree of the node (the number of pedestrians that are in contact with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. (a) corresponds to a simulation without body force and (b) corresponds to a simulation with $k_n=1.2$~E5. The friction coefficient and the other SFM parameters are the same as in Section \ref{bottleneck}. Color online only.}
\label{network_corridor}
\end{figure}

@@

The above results represent an important step in the 
investigation. The inter-pedestrian connectivity differences between the bottleneck geometry 
and the corridor geometry were not expected and opens two major questions:
 how do the pedestrians interact with the walls as
a function of the body stiffness, and consequently, how does 
this affect the flux across the corridor.\\

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_corridor.eps}
\caption{\label{kn_vs_vx_corridor} Mean velocity in the longitudinal coordinate $(v_x)$
 as a function of the stiffness $k_n$ for three different global densities (see label in the plot). 
 The measurements correspond to a corridor of 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$.
 The average was taken along the corridor and along the simulated time. Color online only.      }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./vx_profile.eps}\label{vx_profile}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./strain_rate_vs_kn.eps}\label{strain_rate_vs_kn}}\\
\caption[width=0.47\columnwidth]{(a) Mean velocity in the $x$ coordinate as a function of the transversal coordinate $y$ for three different stiffness values $k_n$ (see the label). (b) Strain rate as a function of the stiffness level $k_n$ for three different global density values (see label). Data correspond to a corridor of 28$\,$m $\times$ 22$\,$m with periodic boundary conditions and $v_d=1$. The global density was $\rho=7 p\,$m$^{-2}$. Color online only. }
\label{profile_strain}
\end{figure}

@@

%It is worth mentioning that the SFM contemplates the possibility of overlapping between pedestrians (unlike the models stated in Refs.~\cite{kabalan_2017,jebrane_2019}). Despite this hard-to-interpret feature, the SFM can reproduce empirical behaviors of collective pedestrian dynamics.\\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./flow-density_multifric_nobodyforce.eps}\label{flow_density_no_body_force}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./flow-density_multifric_bodyforce.eps}\label{flow_density_body_force}}\\
\caption[width=0.47\columnwidth]{Flow vs. global density. The flow is calculated in a circular area of $R=1\,$m at the center of the corridor. The circular markers correspond to the original friction of the SFM, the "+" symbol corresponds to the friction increased by a factor of five and the triangles correspond to the friction increased by a factor of ten. The desired velocity was $v_d=1$ in all the cases. (a) corresponds to simulations without body force ($\mathcal{K}_c =$0) and (b) corresponds to a body force with the original value of the body stiffness ($\mathcal{K}_c =$68). Color online only.}
\label{flow_density}
\end{figure}

>>>>

\begin{table}
\begin{tabular}{c@{\hspace{6mm}}c@{\hspace{6mm}}c@{\hspace{6mm}}c@{\hspace{6mm}}
c@{\hspace{14mm}}l}
 \hline
 $\tau$[s]   & $m$[kg]     & $v_d$[m/s]  &  $B$[m]  & $k_n$[kg/s$^2$] &  Refs.\\
 \hline
0.61         & ---         & 1.24 & $0.36+1.06\,v$ &  ---                 &  
 \cite{seyfried_2007} \\
0.50$^*$     & 80$^*$      & 1.34 & 0.50           &  ---                 &  
\cite{weidmann_1992,lakoba_2005}\\
---          & $67.5$      & 1.39 &  ---           &  $96.1 + 12694.1\,x$ & 
\cite{song_2019}\\
---          & $67.0$      & 1.39 &  ---           &  $97.0 + 29378.9\,x$ & 
\cite{song_2019}\\

@@

\subsection{\label{parameters}The set of parameters}

@@

\subfloat[\label{venn_2parameters}]
{\includegraphics[width=0.49\columnwidth]{./venn_2parameters.png}} \hfill
\subfloat[\label{venn_3parameters}]
{\includegraphics[width=0.49\columnwidth]{./venn_3parameters.png} }\\
\caption[width=0.47\columnwidth]{Venn diagrams for the dimensionless parameters
appearing in the equation of motion (\ref{eqn_motion}) (see
Appendix~\ref{appendix1} for details). The sets correspond to
$\mathcal{A}=\{\tau/m,1/v_d,A\}$, $\mathcal{K}=\{\tau/m,B,k_t\}$ and
$\mathcal{K}_c=\{\tau/m,1/v_d,B,k_n\}$. (a) The Venn diagram representation if
no  body force is introduced in the SFM (only sets $\mathcal{A}$ and
$\mathcal{K}$,  as in Ref.~\cite{dorso_2019}). (b) The Venn diagram
representation for the sets  $\mathcal{A}$, $\mathcal{K}$ and $\mathcal{K}_c$.}
\label{venn_diagram} 
\end{figure}

@@

\begin{itemize}  \item[(a)] The ``faster-is-slower'' (or the ``faster-is-
faster'') phenomenon  occurs when varying the values of $v_d$ (see
Section~\ref{bottleneck}). This is  equivalent to vary simultaneously
$\mathcal{A}$ and $\mathcal{K}_c$ by the same  amount in the (dimensionless)
parameter space (see Section~\ref{parameters}). We  may visualize this sampling
procedure as moving along a straight line in the  space
$(\mathcal{A},\mathcal{K},\mathcal{K}_c)$. Further variations of  $k_n$ turns
the sample points out of this line, but on the plane of constant
$\mathcal{K}_c$. We will step up $k_n$ by several orders of magnitude in order
to get the big picture of this surface.

@@

\item[(c)] We will explore crowd densities allowing body compressions up to
$0.1\,$m. This corresponds to remarkably high densities, since our own  (real-
life) estimates do not surpass $0.065\,$m (see  Section~\ref{experimental}).
Therefore, the extremely high density  scenarios (say, above $0.065\,$m) will be
considered for the purpose of a  tendency, but not as a common real-life
situation. No casualties due to high  pressures will be further considered.

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]{./vd_vs_te_N225.eps}

\caption{\label{vd_vs_te}Mean evacuation time (s) vs. the pedestrian’s desired
velocity (m/s) for a bottleneck. The room was 20 m x 20 m size. The door was
0.92 m width (two pedestrians' width). Mean values were computed from 10
evacuation processes. 225  pedestrians were initially placed in a square lattice
with a random initial  velocity. Each process was finished when 158 pedestrians
left the room. The  different symbols indicate the $k_n$ value corresponding to
the body force (see  the label). The crosses correspond to the Helbing's
original SFM parameter, the  up-triangles correspond to the value measured in
Ref.~\cite{melvin1988aatd}, squares correspond  to zero body force and circles
correspond to an extreme value of stiffness (one  order of magnitude higher than
the original SFM). The   down triangles correspond to an intermediate value
between the empirical value  presented in Ref.~\cite{melvin1988aatd} and the one
provided by Helbing in  Ref.~\cite{helbing_2000} }

\end{figure}

Three behavioral patterns can be distinguished in Fig.~\ref{vd_vs_te}.  Each
pattern can display a positive slope, a negative slope or both. The interval in
which the slope is positive means that the  harder the pedestrians try to get
out (higher $v_d$), the  longer it takes them to evacuate. This is the Faster-
is-slower  (FIS) regime. Conversely, the interval in which the slope is
negative  corresponds to a Faster-is-Faster (FIF) regime (the  harder they try,
the quicker they leave).\\

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the
pedestrian’s desired velocity (m/s). (b) Mean overlap as a function of the
pedestrians desired velocity. Each symbol indicates the $k_n$ value
corresponding to the body force (see the labels). The data corresponds to a
bottleneck with periodic boundary conditions (re-injecting pedestrians). The
average was taken over time and the pedestrians in the bottleneck. The sampling
was done every five seconds once the crowd reached the stationary state (say,
$t=20\,$s) until the end of the simulation ($t=1000\,$s). Color online only.}
\label{degree_overlap_vd}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\

\caption[width=0.47\columnwidth]{Snapshots of the contact networks of a 225
pedestrian evacuation through a bottleneck. The door is placed at
$(x,y)=(20,10)\,$m, the width of the door is 0.92$\,$m (equivalent to 2
pedestrian\textsc{\char13}s diameter). The lines that connect the nodes
(pedestrians) represent the contact between them. The color represents the
degree (the number of pedestrians with which it is connected). (a) and (b)
correspond to a simulation without body force with $v_d=$2 and $v_d=$10
respectively. (c) and (d) correspond to simulations with $k_n=1.2\,$E6 with
$v_d=$2 and $v_d=$10 respectively. Color online only.}
\label{network_bottleneck}
 \end{figure}

@@

It is worth noting from Fig.~\ref{degree_dens} that ``stiff'' pedestrians
exhibit a sharper transition to the maximum packing density than  the ``soft''
pedestrians. We further checked this phenomenon by counting the  number of loops
of three nodes (say, the triangles) present in the network.  This magnitude was
reported to be a sensitive magnitude for characterizing the  jamming transition
in a compressed granular packing (see  Ref.~\cite{pugnaloni_2013} for details).
As can be seen in  Fig.~\ref{triangles}, the ``stiff'' pedestrians share more
triangles  than the ``soft'' ones.\\

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the 
corridor at time $t=50\,$s. The global density was $\rho=6$.  The lines that 
connect the nodes (pedestrians) represent the contact between them. The colors 
stand for the degree of the node (the number of pedestrians that are in contact 
with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary 
conditions and $v_d=1$. (a) corresponds to a simulation without body force and 
(b) corresponds to a simulation with $k_n=1.2$~E5. The friction coefficient and 
the other SFM parameters are the same as in Section \ref{bottleneck}. Color 
online only.}
\label{network_corridor}
\end{figure}

@@

The above results represent an important step in the  investigation. The inter-
pedestrian connectivity differences between the bottleneck geometry  and the
corridor geometry were not expected and opens two major questions:  how do the
pedestrians interact with the walls as a function of the body stiffness, and
consequently, how does  this affect the flux across the corridor.\\

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_corridor.eps}
\caption{\label{kn_vs_vx_corridor} Mean velocity in the longitudinal coordinate
$(v_x)$ as a function of the stiffness $k_n$ for three different global 
densities (see label in the plot). The measurements correspond to a corridor of 
28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. The average 
was taken along the corridor and along the simulated time. Color online only.}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./vx_profile.eps}\label{vx_profile}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./strain_rate_vs_kn.eps}\label{strain_rate_vs_kn}}\\

\caption[width=0.47\columnwidth]{(a) Mean velocity in the $x$ coordinate as a 
function of the transversal coordinate $y$ for three different stiffness values 
$k_n$ (see the label). (b) Strain rate as a function of the stiffness level 
$k_n$ for three different global density values (see label). Data correspond to 
a corridor of 28$\,$m $\times$ 22$\,$m with periodic boundary conditions and 
$v_d=1$. The global density was $\rho=7 p\,$m$^{-2}$. Color online only. }
\label{profile_strain}
\end{figure}

@@

%It is worth mentioning that the SFM contemplates the possibility of overlapping
%between pedestrians (unlike the models stated in
%Refs.~\cite{kabalan_2017,jebrane_2019}). Despite this hard-to-interpret
%feature, the SFM can reproduce empirical behaviors of collective pedestrian
%dynamics.\\

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./flow-density_multifric_nobodyforce.eps}\label{flow_density_no_body_force}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./flow-density_multifric_bodyforce.eps}\label{flow_density_body_force}}\\
\caption[width=0.47\columnwidth]{Flow vs. global density. The flow is calculated
in a circular area of $R=1\,$m at the center of the corridor. The circular 
markers correspond to the original friction of the SFM, the "+" symbol 
corresponds to the friction increased by a factor of five and the triangles 
correspond to the friction increased by a factor of ten. The desired velocity 
was $v_d=1$ in all the cases. (a) corresponds to simulations without body force 
($\mathcal{K}_c =$0) and (b) corresponds to a body force with the original value
 of the body stiffness ($\mathcal{K}_c =$68). Color online only.}
\label{flow_density}
\end{figure}
